---
title: "Industrial Load Simulation"
author: "Ian Bledsoe"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
knit: (
  function(inputFile, encoding) { 
  
    rmarkdown::render( 
      input       = inputFile, 
      encoding    = encoding, 
      output_file = paste(Sys.Date(), 'Industrial Load Sim')) })
---

<!---
Hello World! (MIDS application Reviewers)
This R script is a Monte-Carlo Simulation I wrote.
It takes the most recent load forecast, which is a point forecast using a linear regression model
Then it creates 800 random futures based on a normal distribution with mean equal to the point forecast
and standard deviation equal to the SD of historic normalized loads.
Also simulates a behind the meter resource (Cogen) which is covariant with one of the loads (Wauna).

Please excuse the mess! =)
-->

<!---
Tips:
  Don't name code chunks 'setup' - reserved
  include=false means you wont see any output
  echo=false means you wont see code, but will see output of code?
  need 1 space after # for headings/subheadings
  
  May fail knitting to .html?
  worked to Word.
  
  Include - Display output (if any) of code in doc
  Echo - Display code in output doc
  Eval - code doesnt run at all
-->


<!-- DT (datatable) is a cool package for displaying tables in .rmd -->
<!-- "esquisse" looks cool for GUI graphics -->

# Purpose of this script is to generate loads for 800 simulated futures
Camas, Halsey, Wauna&Cogen - wauna & cogen should be covariant.
Used in the DPM.
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```


#Packages
```{r loadpkgs, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, results='hide'}
library(lubridate)#date manipulation library *floor_date
library(svMisc)
library(magrittr) #pipe operators
library(reshape2) #pivot
library(crayon) #allows print (text) with color
library(gridExtra) #need this for display of lattice graphs
library(MASS)#for mvnorm, scale
library(marima)
library(ggplot2)
library(MLmetrics) #MAPE
library(RODBC)
library(ggpubr)
library(grid)
library(dplyr) #load last; plyr conflict (what abt other packages that import plyr?)
cat("\014") #clear console

result = tryCatch({
    dev.off()
}, error = function(error_condition) {
  })
rm(list=setdiff(ls(), c(""))) #remove objects
source('\\\\itv00005\\Power\\Data_Systems\\R\\CustomFunctions.R')
```


```{r sql, echo=FALSE}
channel <- odbcConnect("Superpower")
rawload<-sqlQuery(channel, "execute Meter_Mathv2") #Slow ~20s
loadhour<-sqlQuery(channel, "execute LAT_LoadHour_Vertical")
outdata<-sqlQuery(channel, "execute LAT_Outages") #Slow ~30s
#factordata<-sqlQuery(channel, "execute LAT_Load_Factors") #Slow ~15s
raweedata<-sqlQuery(channel, "execute LAT_Energy_Efficiency") #Slow ~15s
forecastdata<-sqlQuery(channel, "Select atable.*, LH as [Hrs]
                                  from
                                  (
                                  select * from Load_Forecast
                                  where Run_Date = (select Max(Run_Date) from Load_Forecast)
                                  )atable
                                  left join
                                  (
                                  select cast(DATEADD(month, DATEDIFF(month, 0, [Timestamp]), 0)AS date) as Dt
                                  		,Load_Hour
                                  		,count(Load_Hour) as LH	
                                  from Workday_Timeseries
                                  group by cast(DATEADD(month, DATEDIFF(month, 0, [Timestamp]), 0)AS date)
                                  			,Load_Hour
                                  )btable
                                  on atable.MonthYear=btable.Dt
                                  and atable.Load_Hour=btable.Load_Hour")
close(channel); rm(channel)
```

```{r clean, echo=FALSE}

loaddata<- rawload %>%
                #replace(., is.na(.), 0) %>%
                group_by(Dy=as.Date(floor_date(Timestamp, "month"))) %>%
                summarise(Wauna=mean(na.omit(Wauna)),
                          Cogen=mean(na.omit(WaunaCoGen)),
                          Halsey=mean(na.omit(Halsey)),
                          Camas=mean(na.omit(Camas)))
loaddata<-loaddata[loaddata$Dy>="2004-01-01" & loaddata$Dy < as.Date(cut(Sys.Date(), "month")),]
loaddata$Month<-month(loaddata$Dy)
loaddata$timeseq<-0:(nrow(loaddata)-1)
loaddata$revtimeseq<-rev(0:(nrow(loaddata)-1))
#EE Data
raweedata$Month_Claimed<-as.Date(raweedata$Month_Claimed)
mframe<-data.frame("MonYr"=as.Date(seq(ymd(min(raweedata$Month_Claimed)),ymd(floor_date(Sys.Date(), "month") - months(1)), by = 'months')))
eedata<-merge(mframe, raweedata, by.x="MonYr", by.y="Month_Claimed", all=T)
eedata[is.na(eedata)]<-0
eedata<-merge(eedata, loaddata, by.x="MonYr", by.y="Dy", all.x=T)
eedata$wauna.ee.pct<-eedata$wauna.ee/eedata$Wauna
eedata$halsey.ee.pct<-eedata$halsey.ee/eedata$Halsey
##eedata<-eedata[eedata$MonYr>='2004-01-01' & eedata$MonYr<'2019-01-01' ,]
eedata$cumWauna<-as.numeric(rev(cumsum(rev(eedata$wauna.ee))))
eedata$cumHalsey<-as.numeric(rev(cumsum(rev(eedata$halsey.ee))))
loaddata<-merge(loaddata, eedata[,c(1, 15, 16)], by.x="Dy", by.y="MonYr", all.x=T)
#Out Data
outdata$Dy<-as.Date(outdata$Dy)
outdata <- outdata %>%
              group_by(Dy) %>%
              summarise(WaunaOut=sum(WaunaOut),
                        HalseyOut=sum(HalseyOut),
                        CamasOut=sum(CamasOut),
                        CogenOut=sum(CogenOut))
loaddata<-merge(loaddata, outdata, by.x="Dy", by.y="Dy", all.x=T)
#ForecastData

forecastdata<-forecastdata[,c(1,4,6,8)] %>%
                  group_by(MonthYear,Variable) %>%
                  summarise(MWh=sum(Quantity),
                            Hrs=sum(Hrs)) %>%
                  mutate(aMW=MWh/Hrs,
                         MWh=NULL,
                         Hrs=NULL) %>%
                  dcast(MonthYear ~ Variable, value.var="aMW")
forecastdata$MonthYear<-as.Date(forecastdata$MonthYear)
```


#Outage normalization
```{r outage_norm2, echo=FALSE, include=TRUE}
waunaoutmod<-lm(Wauna~WaunaOut, data=loaddata[!(loaddata$Month == 5 & loaddata$Dy > '2010-01-01'),])
loaddata$Wauna.norm1<-loaddata$Wauna
loaddata[!(loaddata$Month == 5 & loaddata$Dy > '2010-01-01'),]$Wauna.norm1<-as.numeric(loaddata[!(loaddata$Month == 5 & loaddata$Dy > '2010-01-01'),]$Wauna - (waunaoutmod$coefficients[2] * loaddata[!(loaddata$Month == 5 & loaddata$Dy > '2010-01-01'),]$WaunaOut))

halseyoutmod<-lm(Halsey~HalseyOut, data=loaddata)
loaddata$Halsey.norm1<-loaddata$Halsey - (halseyoutmod$coefficients[2] * loaddata$HalseyOut)

cogenoutmod<-lm(Cogen~CogenOut, data=loaddata[!(loaddata$Month == 5 & loaddata$Dy > '2010-01-01'),])
loaddata$Cogen.norm1<-loaddata$Cogen
loaddata[!(loaddata$Month == 5 & loaddata$Dy > '2010-01-01'),]$Cogen.norm1<-as.numeric(loaddata[!(loaddata$Month == 5 & loaddata$Dy > '2010-01-01'),]$Cogen - (cogenoutmod$coefficients[2] * loaddata[!(loaddata$Month == 5 & loaddata$Dy > '2010-01-01'),]$CogenOut))


```

#Remove EE
```{r remove_ee, echo=FALSE, include=FALSE}
#Scale EE to Month
ratioframe <- loaddata %>%
              group_by(Month=month(Dy)) %>%
              summarise(WaunaRatio=mean(Wauna),
                        HalseyRatio=mean(Halsey))
ratioframe$WaunaRatio<-ratioframe$WaunaRatio/mean(loaddata$Wauna)
ratioframe$HalseyRatio<-ratioframe$HalseyRatio/mean(loaddata$Halsey)
loaddata<-merge(loaddata, ratioframe, by.x="Month", by.y="Month", all.x=T)
loaddata<-loaddata[order(loaddata$Dy),]
loaddata$cumWauna<-loaddata$cumWauna * loaddata$WaunaRatio
loaddata$cumHalsey<-loaddata$cumHalsey * loaddata$HalseyRatio
#Remove historic EE
loaddata$Wauna.norm2<-loaddata$Wauna.norm - loaddata$cumWauna
loaddata$Halsey.norm2<-loaddata$Halsey.norm - loaddata$cumHalsey
loaddata<-loaddata[order(loaddata$Dy),]
```

#Detrending
```{r detrend, echo=TRUE, include=FALSE}
#Halsey
halseymod<-lm(Halsey.norm2~timeseq, data=loaddata, na.action = na.omit)
loaddata$Halsey.norm3<-na.omit(loaddata$Halsey.norm2) + (halseymod$coefficients[2] * na.omit(loaddata$revtimeseq))

#Detrend Wauna
waunamod<-lm(Wauna.norm2~timeseq, data=loaddata)
loaddata$Wauna.norm3<-loaddata$Wauna.norm2 + (waunamod$coefficients[2] * loaddata$revtimeseq)

#Detrend Cogen
cogenmod<-lm(Cogen.norm1~timeseq, data=loaddata)
loaddata$Cogen.norm3<-loaddata$Cogen.norm1 + (cogenmod$coefficients[2] * loaddata$revtimeseq)

#Detrend Camas
camasframe<-loaddata[loaddata$Dy<"2018-09-01",]
camasframe<-camasframe[complete.cases(camasframe[, 6]),]
camasframe$timeseq<-seq(0,NROW(camasframe)-1)
camasframe$revtimeseq<-rev(seq(0,NROW(camasframe)-1))
camasmod<-lm(Camas~timeseq, data=camasframe, na.action=na.omit)
loaddata$Camas.norm3<-loaddata$Camas
loaddata[loaddata$Dy<"2018-09-01" & complete.cases(loaddata[, 6]),]$Camas.norm3<-camasframe$Camas + (camasmod$coefficients[2] * camasframe$revtimeseq)

loaddata[loaddata$Cogen.norm3<0,]$Cogen.norm3<-0
```

#Simulation
```{r simulation, echo=F}

sdincrease<-1.35

loaddata<-loaddata[order(loaddata$Dy),]

simframe<-data.frame("MonYr"=rep(as.Date(seq(ymd("2019-01-01"),ymd("2040-12-1"), by = 'months')), times=800),
                     "Run"=rep(1:800, each=(12*22)),
                     "Month"=month(rep(as.Date(seq(ymd("2019-01-01"),ymd("2040-12-1"), by = 'months')), times=800)),
                     "Wauna"= NA,
                     "Cogen"= NA,
                     "Halsey"= NA,
                     "Camas"= NA)

#Sim Halsey
for (yr_no in 2019:2040){
    for (mon_no in 1:12){
      simframe[simframe$Month==mon_no & year(simframe$MonYr)==yr_no,]$Halsey<-rnorm(NROW(simframe[month(simframe$MonYr)==mon_no & year(simframe$MonYr)==yr_no,]),mean=mean(forecastdata[month(forecastdata$MonthYear)==mon_no & year(forecastdata$MonthYear)==yr_no,]$Halsey),
                                                      sd=sd(loaddata$Halsey.norm3)*sdincrease)
    }
}

#sim camas, build monthly distribution from hourly data...? help deal with short dataset?
#SD is not month-specific, since you need at least two data.
for (yr_no in 2019:2040){
    for (mon_no in 1:12){
      simframe[simframe$Month==mon_no & year(simframe$MonYr)==yr_no,]$Camas<-rnorm(NROW(simframe[month(simframe$MonYr)==mon_no & year(simframe$MonYr)==yr_no,]), mean= mean(forecastdata[month(forecastdata$MonthYear)==mon_no & year(forecastdata$MonthYear)==yr_no,]$Camas), 
sd=sd(na.omit(loaddata[loaddata$Dy>="2018-09-01",]$Camas.norm3))*sdincrease) #works, but extremes not extreme enough
#sd=sd(na.omit(loaddata[loaddata$Month==mon_no,]$Camas.norm3))) #too much (impossible) variation...
    }
}

#Sim covariant Wauna & Cogen
for (yr_no in 2019:2040){
  for (mon_no in 1:12){
    covmx<-cov(loaddata[loaddata$Month==mon_no,c("Wauna.norm3", "Cogen.norm3")])
    covmx<-covmx*sdincrease
    means<-c(mean(forecastdata[month(forecastdata$MonthYear)==mon_no & year(forecastdata$MonthYear)==yr_no,]$Wauna)
            ,mean(loaddata[loaddata$Month==mon_no & yr_no%%3==year(loaddata$Dy)%%3,]$Cogen.norm3)) # 
    draws <- mvrnorm(NROW(simframe[month(simframe$MonYr)==mon_no & year(simframe$MonYr)==yr_no,]),
                   mu=means,
                  Sigma=covmx)
  simframe[simframe$Month==mon_no & year(simframe$MonYr)==yr_no,]$Wauna<-draws[,"Wauna.norm3"]
  simframe[simframe$Month==mon_no & year(simframe$MonYr)==yr_no,]$Cogen<-draws[,"Cogen.norm3"]
  }
}

simframe[simframe$Cogen <0,]$Cogen <- 0
simframe$Year<-year(simframe$MonYr)
```

#Summary Frames
```{r create_megaframe, echo=FALSE}

loaddata$Type<-"Actual"
loaddata$Run<-NA
simframe$Type<-"Simulated"
forecastdata$Type<-"Forecast"
forecastdata$Run<-NA
forecastdata$Cogen<-NA


#Does this work, or does it blend actual, sim & forecast????
megaframe<-as.data.frame(mapply(c, loaddata[, c("Dy", "Halsey.norm3", "Wauna.norm3", "Cogen.norm3", "Camas.norm3", "Type", "Run")],
                                   simframe[,c("MonYr", "Halsey", "Wauna", "Cogen", "Camas", "Type", "Run")],
                                    forecastdata[,c("MonthYear", "Halsey", "Wauna", "Cogen", "Camas", "Type", "Run")], SIMPLIFY = F))

megaframe<-megaframe %>% 
            rename(
              Halsey=Halsey.norm3,
              Wauna=Wauna.norm3,
              Cogen=Cogen.norm3,
              Camas=Camas.norm3
              )

megaframe$Year<-year(megaframe$Dy)
megaframe[is.na(megaframe$Camas),]$Camas <- mean(na.omit(forecastdata$Camas))
megaframe$Total<-megaframe$Halsey+megaframe$Wauna+megaframe$Camas
megaframe$TotalexCamas<-megaframe$Halsey+megaframe$Wauna

yrframe<-megaframe %>%
        group_by(Year=year(Dy),
                 Run=Run,
                 Type=Type) %>%
        summarise(Halsey=weighted.mean(Halsey,days_in_month(Dy)),
                  Wauna=weighted.mean(Wauna,days_in_month(Dy)),
                  Cogen=weighted.mean(Cogen,days_in_month(Dy)),
                  Camas=weighted.mean(Camas,days_in_month(Dy)))


yrframe$Total<- yrframe$Halsey+yrframe$Wauna+yrframe$Camas
yrframe$TotalexCamas<- yrframe$Halsey+yrframe$Wauna



```

# Histograms!

```{r plot6, echo=FALSE}

#plot monthly histos old vs new. facet!

#Wauna
ggplot(megaframe, aes(Wauna)) +
    geom_histogram(aes(fill=Type),show.legend=FALSE, binwidth=1) +
    facet_grid(Type~., scales="free") +
    ggtitle("Wauna Monthly Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=seq(0, max(na.omit(megaframe$Wauna)), 1))

#Halsey
ggplot(megaframe, aes(Halsey)) +
    geom_histogram(aes(fill=Type),show.legend=FALSE, binwidth=1) +
    facet_grid(Type~., scales="free") +
    ggtitle("Halsey Monthly Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=seq(0, max(na.omit(megaframe$Halsey)), 1))

#Camas
ggplot(megaframe, aes(Camas)) +
    geom_histogram(aes(fill=Type),show.legend=FALSE, binwidth=1) +
    facet_grid(Type~., scales="free") +
    ggtitle("Camas Monthly Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=seq(0, max(na.omit(megaframe$Camas)), 1))

#Cogen
ggplot(megaframe[!is.na(megaframe$Cogen),], aes(Cogen)) +
    geom_histogram(aes(fill=Type),show.legend=FALSE, binwidth=1) +
    facet_grid(Type~., scales="free") +
    ggtitle("Cogen Monthly Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=seq(0, max(na.omit(megaframe$Cogen)), 2))







probs <- c(0.05, 0.95)
xlims<- c(80, 116)

######################################

#monthly actual dist setup
actmonframe<- megaframe[megaframe$Type=="Actual",]
actmondens<- density(actmonframe$TotalexCamas) #########excludes camas
actmondensframe<-data.frame(x=actmondens$x, y=actmondens$y)
actmonquantiles <- quantile(actmonframe$TotalexCamas, prob=probs) #########excludes camas
actmondensframe$quant <- factor(findInterval(actmondensframe$x,actmonquantiles))

#monthly sim dist setup (2019)
simmonframe<- megaframe[megaframe$Type=="Simulated" & megaframe$Year==2019,]
simmondens<- density(simmonframe$TotalexCamas) #########excludes camas
simmondensframe<-data.frame(x=simmondens$x, y=simmondens$y)
simmonquantiles <- quantile(simmonframe$TotalexCamas, prob=probs) #########excludes camas
simmondensframe$quant <- factor(findInterval(simmondensframe$x,simmonquantiles))


cat(blue("Total Monthly Quantiles and 90% bands: "));cat("\n")
cat("Sim p5: ", simmonquantiles[[1]], "\n")
cat("Sim p95: ", simmonquantiles[[2]], "\n")
cat("Sim Spread: ", simmonquantiles[[2]]-simmonquantiles[[1]], "\n\n")

cat("Actual p5: ", actmonquantiles[[1]], "\n")
cat("Actual p95: ", actmonquantiles[[2]], "\n")
cat("Actual Spread: ", actmonquantiles[[2]]-actmonquantiles[[1]], "\n\n")

simmondist <- ggplot(simmondensframe, aes(x,y)) +
            geom_line() +
            geom_ribbon(aes(ymin=0, ymax=y, fill=quant)) +
            scale_x_continuous(breaks=all_int_breaks(yrframe$Total), limits=xlims) +
            scale_fill_brewer(guide="none") +
            ggtitle("Simulated Theoretical Monthly Distribution (2019)")

actmondist <- ggplot(actmondensframe, aes(x,y)) +
            geom_line() +
            geom_ribbon(aes(ymin=0, ymax=y, fill=quant)) +
            scale_x_continuous(breaks=all_int_breaks(yrframe$Total), limits=xlims) +
            scale_fill_brewer(guide="none") +
            ggtitle("Actual Theoretical Monthly Distribution")



ggarrange(actmondist,simmondist, align = "hv", ncol = 1, nrow = 4)

```


```{r plot7, echo=FALSE}
#plot annual histos old vs new. facet!

#Wauna
ggplot(yrframe, aes(Wauna)) +
    geom_histogram(aes(fill=Type),show.legend=FALSE, binwidth=1) +
    facet_grid(Type~., scales="free") +
    ggtitle("Wauna Annual Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=all_int_breaks(yrframe$Wauna))

#Halsey
ggplot(yrframe, aes(Halsey)) +
    geom_histogram(aes(fill=Type),show.legend=FALSE, binwidth=1) +
    facet_grid(Type~., scales="free") +
    ggtitle("Halsey Annual Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=all_int_breaks(yrframe$Halsey))

#Camas
ggplot(yrframe, aes(Camas)) +
    geom_histogram(aes(fill=Type),show.legend=FALSE, binwidth=1) +
    facet_grid(Type~., scales="free") +
    ggtitle("Camas Annual Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=all_int_breaks(na.omit(yrframe$Camas)))

#Cogen
ggplot(yrframe[!is.na(yrframe$Cogen),], aes(Cogen)) +
    geom_histogram(aes(fill=Type),show.legend=FALSE, binwidth=1) +
    facet_grid(Type~., scales="free") +
    ggtitle("Cogen Annual Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=all_int_breaks(na.omit(yrframe$Cogen)))


#Total
ggplot(yrframe, aes(Total)) +
    geom_histogram(aes(fill=Type),show.legend=FALSE, binwidth=1) +
    facet_grid(Type~., scales="free") +
    ggtitle("Total Annual Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=all_int_breaks(yrframe$Total))

#annual histograms by year
ggplot(yrframe[yrframe$Type=="Simulated",], aes(Total)) +
    geom_histogram(aes(fill=Year),binwidth=1) +
    facet_grid(Year~., scales="free") +
    ggtitle("Total Annual Sim Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=all_int_breaks(yrframe$Total))




#annual histograms by year
ggplot(yrframe[yrframe$Type=="Actual",], aes(Total)) +
    geom_histogram(aes(fill=Year),binwidth=1) +
    facet_grid(Year~., scales="free") +
    ggtitle("Total Annual Actual Histograms") +
    scale_x_continuous(minor_breaks=NULL, breaks=all_int_breaks(yrframe$Total))


```


#Probability Distributions

```{r percentile_yr}

probs <- c(0.05, 0.95)
xlims<- c(95, 116)

######################################

#annual actual hist setup
actyrframe<- yrframe[yrframe$Type=="Actual",]
actdens<- density(actyrframe$Total)
actdensframe<-data.frame(x=actdens$x, y=actdens$y)
actquantiles <- quantile(actyrframe$Total, prob=probs)
actdensframe$quant <- factor(findInterval(actdensframe$x,actquantiles))

#annual sim hist setup (2019)
simyrframe<- yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]
simdens<- density(simyrframe$Total)
simdensframe<-data.frame(x=simdens$x, y=simdens$y)
simquantiles <- quantile(simyrframe$Total, prob=probs)
simdensframe$quant <- factor(findInterval(simdensframe$x,simquantiles))

#######################################

#annual actual hist setup (No Camas)
actdens2<- density(actyrframe$TotalexCamas)
actdensframe2<-data.frame(x=actdens2$x, y=actdens2$y)
actquantiles2 <- quantile(actyrframe$TotalexCamas, prob=probs)


#annual sim hist setup (2019) (No Camas)
simdens2<- density(simyrframe$TotalexCamas)
simdensframe2<-data.frame(x=simdens2$x, y=simdens2$y)
simquantiles2 <- quantile(simyrframe$TotalexCamas, prob=probs)


######################################

simdist <- ggplot(simdensframe, aes(x,y)) +
            geom_line() +
            geom_ribbon(aes(ymin=0, ymax=y, fill=quant)) +
            scale_x_continuous(breaks=all_int_breaks(yrframe$Total), limits=xlims) +
            scale_fill_brewer(guide="none") +
            ggtitle("Simulated Theoretical Distribution (2019)")

#annual sim histogram (2020)
simhist <- ggplot(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,], aes(Total)) +
            geom_histogram(binwidth=1) +
            ggtitle("Simulated Histogram (2019)") +
            scale_x_continuous(minor_breaks=NULL, breaks=all_int_breaks(yrframe$Total), limits=xlims)

actdist <- ggplot(actdensframe, aes(x,y)) +
            geom_line() +
            geom_ribbon(aes(ymin=0, ymax=y, fill=quant)) +
            scale_x_continuous(breaks=all_int_breaks(yrframe$Total), limits=xlims) +
            scale_fill_brewer(guide="none") +
            ggtitle("Actual Normalized Theoretical Distribution")

acthist <- ggplot(yrframe[yrframe$Type=="Actual",], aes(Total)) +
            geom_histogram(aes(fill=Year),binwidth=1) +
            ggtitle("Actual Normalized Histogram") +
            scale_x_continuous(minor_breaks=NULL, breaks=all_int_breaks(yrframe$Total), limits=xlims)

ggarrange(simdist,simhist,actdist,acthist, align = "hv", ncol = 1, nrow = 4)

#Error about missing rows seems to be a lie
```

#Summary Stats

```{r sum_stats}
#plot summary stats

cat(red("Hist. normalized Wauna: \n"));summary(loaddata$Wauna.norm3); cat(green("Random Wauna: \n"));summary(simframe[simframe$Year==2019,]$Wauna); cat(black("***************\n"))
cat(red("Hist. normalized Cogen: \n"));summary(loaddata$Cogen.norm3); cat(green("Random Cogen: \n"));summary(simframe[simframe$Year==2019,]$Cogen); cat(black("***************\n"))
cat(red("Hist. normalized Halsey: \n"));summary(loaddata$Halsey.norm3); cat(green("Random Halsey: \n"));summary(simframe[simframe$Year==2019,]$Halsey); cat(black("***************\n"))
cat(red("Hist. normalized Camas: \n"));summary(loaddata$Camas.norm3); cat(green("Random Camas: \n"));summary(simframe[simframe$Year==2019,]$Camas); cat(black("***************\n"))


# Makes sense for random to be lower, since they're based off
# forecast, which is decreasing as time goes on.
```


```{r sum_stats2}
cat(blue("Annual Stats\n"))

cat(red("Actual.n Total Mean: ")); mean(yrframe[yrframe$Type=="Actual",]$Total)
cat(red("Sim Total mean (2019): ")); mean(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]$Total); cat("\n")

cat(red("Actual.n Total SD: ")); sd(yrframe[yrframe$Type=="Actual",]$Total)
cat(red("Sim Total SD (2019): ")); sd(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]$Total); cat("\n")

cat(red("Act.n Halsey SD: ")); sd(yrframe[yrframe$Type=="Actual",]$Halsey)
cat(red("Sim Halsey SD (2019): ")); sd(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]$Halsey); cat("\n")

cat(red("Act.n Wauna SD: ")); sd(yrframe[yrframe$Type=="Actual",]$Wauna)
cat(red("Sim Wauna SD (2019): ")); sd(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]$Wauna); cat("\n")

cat(red("Act.n Cogen SD: ")); sd(yrframe[yrframe$Type=="Actual",]$Cogen)
cat(red("Sim Cogen SD (2019): ")); sd(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]$Cogen); cat("\n")

cat(red("Act.n Camas SD: ")); sd(na.omit(yrframe[yrframe$Type=="Actual",]$Camas))
cat(red("Sim Camas SD (2019): ")); sd(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]$Camas); cat("\n")


cat(blue("Total Annual Quantiles and 90% bands: "));cat("\n")
cat("Sim p5: ", simquantiles[[1]], "\n")
cat("Sim p95: ", simquantiles[[2]], "\n")
cat("Sim Spread: ", simquantiles[[2]]-simquantiles[[1]], "\n\n")
cat("Actual p5: ", actquantiles[[1]], "\n")
cat("Actual Spread: ", actquantiles[[2]]-actquantiles[[1]], "\n\n")

```


```{r sum_stats2}
cat(blue("Monthly Standard Deviations\n"))

cat(red("Actual.n Total SD: ")); sd(simframe[simframe$Type=="Actual",]$Total)
cat(red("Sim Total SD: ")); sd(simframe[simframe$Type=="Simulated" & simframe$Year==2019,]$Total); cat("\n")

cat(red("Act.n Halsey SD: ")); sd(yrframe[yrframe$Type=="Actual",]$Halsey)
cat(red("Sim Halsey SD: ")); sd(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]$Halsey); cat("\n")

cat(red("Act.n Wauna SD: ")); sd(yrframe[yrframe$Type=="Actual",]$Wauna)
cat(red("Sim Wauna SD: ")); sd(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]$Wauna); cat("\n")

cat(red("Act.n Wauna SD: ")); sd(yrframe[yrframe$Type=="Actual",]$Cogen)
cat(red("Sim Wauna SD: ")); sd(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]$Cogen); cat("\n")

cat(red("Act.n Camas SD: ")); sd(na.omit(yrframe[yrframe$Type=="Actual",]$Camas))
cat(red("Sim Camas SD: ")); sd(yrframe[yrframe$Type=="Simulated" & yrframe$Year==2019,]$Camas); cat("\n")

```

# 2d Bin Plots

```{r plot4, echo=FALSE}

#Wauna
ggplot(simframe, aes(MonYr, Wauna)) +
    geom_bin2d() +
    geom_line(data=forecastdata, aes(MonthYear, Wauna), size=1, col="red")
    #geom_ribbon(aes(min=min(Wauna), max=max(Wauna)))

#Halsey
ggplot(simframe, aes(MonYr, Halsey)) +
    geom_bin2d() +
    geom_line(data=forecastdata, aes(MonthYear, Halsey), size=1, col="red")
    #geom_ribbon(aes(min=min(Wauna), max=max(Wauna)))

#Camas
ggplot(simframe, aes(MonYr, Camas)) +
    geom_bin2d() +
    geom_line(data=forecastdata, aes(MonthYear, Camas), size=1, col="red")
    
#Cogen
ggplot(simframe, aes(MonYr, Cogen)) +
    geom_bin2d()

```


# Plot Outage Normalized 
Actuals (black) & outage-normalized (red)
```{r plot1, echo=FALSE, include=TRUE}
lnwght=1
#Wauna
ggplot(loaddata, aes(Dy, Wauna)) +
    geom_line(size=lnwght) +
    geom_line(aes(Dy, Wauna.norm1),size=lnwght+1,col="red")

#Cogen
ggplot(loaddata, aes(Dy, Cogen)) +
    geom_line(size=lnwght) +
    geom_line(aes(Dy, Cogen.norm1),size=lnwght+1,col="red")

#Halsey
ggplot(loaddata[!is.na(loaddata$Halsey),], aes(Dy, Halsey)) +
  geom_line(size=lnwght) +
  geom_line(aes(Dy, Halsey.norm1),col="red", size=lnwght+1)

#Camas
ggplot(loaddata[!is.na(loaddata$Camas),], aes(Dy, Camas)) +
  geom_line(size=lnwght)

```



# Plot EE removed
Actuals (black), outage normalized (red), and outage normalized with ee removed (blue)
```{r plot2, echo=FALSE}
lnwght=1

#Wauna
ggplot(loaddata, aes(Dy, Wauna)) +
    geom_line(size=lnwght) +
    geom_line(aes(Dy, Wauna.norm1),size=lnwght,col="red") +
    geom_line(aes(Dy, Wauna.norm2),size=lnwght+1,col="blue")

#Cogen
ggplot(loaddata, aes(Dy, Cogen)) +
    geom_line(size=lnwght) +
    geom_line(aes(Dy, Cogen.norm1),size=lnwght,col="red")

#Halsey
ggplot(loaddata[!is.na(loaddata$Halsey),], aes(Dy, Halsey)) +
  geom_line(size=lnwght) +
  geom_line(aes(Dy, Halsey.norm1),col="red", size=lnwght) +
  geom_line(aes(Dy, Halsey.norm2),col="blue", size=lnwght+1)

#Camas
ggplot(loaddata[!is.na(loaddata$Camas),], aes(Dy, Camas)) +
  geom_line(size=lnwght)

```

# Plot Trend Removed
Actuals (black), outage normalized (red), outage normalized with ee removed (blue), and outage normalized with ee removed and trend removed (green)

```{r plot3, echo=FALSE,}
lnwght=1

#Wauna
ggplot(loaddata, aes(Dy, Wauna)) +
    geom_line(size=lnwght) +
    geom_line(aes(Dy, Wauna.norm1),size=lnwght,col="red") +
    geom_line(aes(Dy, Wauna.norm2),size=lnwght,col="blue") +
    geom_line(aes(Dy, Wauna.norm3),size=lnwght+1,col="green4")

#Cogen
ggplot(loaddata, aes(Dy, Cogen)) +
    geom_line(size=lnwght) +
    geom_line(aes(Dy, Cogen.norm1),size=lnwght,col="red")+
    geom_line(aes(Dy, Cogen.norm3),size=lnwght+1,col="green4")

#Halsey
ggplot(loaddata[!is.na(loaddata$Halsey),], aes(Dy, Halsey)) +
  geom_line(size=lnwght) +
  geom_line(aes(Dy, Halsey.norm1),col="red", size=lnwght) +
  geom_line(aes(Dy, Halsey.norm2),col="blue", size=lnwght)+
  geom_line(aes(Dy, Halsey.norm3),col="green4", size=lnwght+1)

#Camas
ggplot(loaddata[!is.na(loaddata$Camas),], aes(Dy, Camas)) +
  geom_line(size=lnwght) +
  geom_line(aes(Dy, Camas.norm3),col="green4", size=lnwght+1)

```

```{r somestuff, include=F}
summary(megaframe[megaframe$Type=="Actual",]);cat("\n")
summary(megaframe[megaframe$Type=="Simulated",]);cat("\n")
summary(megaframe[megaframe$Type=="Forecast",])

#Proportion of Cogen zeros Actual
length(which(megaframe[megaframe$Type=="Actual",]$Cogen == 0))/nrow(megaframe[megaframe$Type=="Actual",])

#Proportion of cogen zeros - Simulated
length(which(megaframe[megaframe$Type=="Simulated",]$Cogen == 0))/nrow(megaframe[megaframe$Type=="Simulated",])

#Are there actually zeroes for Camas?
length(which(megaframe[megaframe$Type=="Simulated",]$Camas == 0))
length(which(megaframe[megaframe$Type=="Actual",]$Camas == 0))
length(which(megaframe[megaframe$Type=="Simulated",]$Camas == 0))

```

```{r outputmelt, echo=F}
outputframe<-simframe %>%
        melt(id.vars=c("MonYr",
                       "Run",
                       "Type"),
             measure.vars = c("Wauna",
                              "Cogen",
                              "Halsey",
                              "Camas")) %>%
        transmute(Source= "PwrAnalyst", 
                  PublishDate= as.Date(Sys.time()), 
                  Description=Type ,
                  Year= year(MonYr),
                  Quarter= quarter(MonYr),
                  Period_Start= MonYr,
                  Period_End= ceiling_date(MonYr, "month") - days(1) ,
                  Run=Run,
                  Variable= variable,
                  Unit = "aMW",
                  Value=value )
```

```{r testx}
mean(actyrframe[actyrframe$

```



```{r writeoutput, include=F}
  write.csv(outputframe, file= (paste0("\\\\itv00006\\sqlupload$", "\\Simulation_Data.csv")),row.names=FALSE, quote=FALSE)
```